---
title: "Data Visualization"
author: "Guillermo Romero"
date: "2022-08-17"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
```

\# Read in Lizard Data

```{r}
lizards <- read_csv(here::here("data_tidy","lizards.csv"))
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + geom_point()


```

create histogram for total lengths in lizards data set

```{r}
ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.3,
                 linetype = "dotted") 
# if you are not changing a variable on x and y then don't put inside aes
  
```

```{r}
ggplot(data = lizards, aes( x = total_length, y = weight)) +
  geom_point(shape = 22,
             color = "cyan4",
             fill = "yellow",
             size = 4,
             alpha = 0.4)
```

changing the aesthic based on different values of a variable

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length), alpha = 0.5) +
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() + facet_wrap(~common_name, ncol = 3, scales = "free")
```

```{r}
lizards_counts <- lizards |> 
  group_by(common_name) |> 
  summarize(counts = n())
lizards_counts

liz_count <- lizards |> 
  count(common_name, site, tail)
liz_count
```

```{r}
ggplot(lizards_counts, aes(y = common_name, x = counts)) +
  geom_col() # default for categorical variable is alphabetical variable, can convert to factor and set the order
# convert and plot as factor by the value of n
# manual orders are not reproducible 
```

# Factor reorder

```{r}
ggplot(lizards_counts, aes(y = fct_reorder(common_name, counts), x = counts)) +
  geom_col()
```

Create scatterplots, faceted by species, for total length versus weight.

The color and shape of points should differ based on the species name (common name). Change theme. Make each faceted panel have flexible axis limits (scales should be free).

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name),
             alpha = 0.6,
             size = 3) +
  facet_wrap(~common_name, scales = "free") +
  theme_light() +
  labs(x = " Total Length (mm)",
       y = "Weight (grams)")
```
