---
title: "Data Visualization"
author: "Guillermo Romero"
date: "2022-08-17"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(here)
library(janitor)
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)
library(gapminder)
library(sf)
library(ggthemes)
```

\# Read in Lizard Data

```{r}
lizards <- read_csv(here::here("data_tidy","lizards.csv"))
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) + geom_point()


```

create histogram for total lengths in lizards data set

```{r}
ggplot(data = lizards, aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.3,
                 linetype = "dotted") 
# if you are not changing a variable on x and y then don't put inside aes
  
```

```{r}
ggplot(data = lizards, aes( x = total_length, y = weight)) +
  geom_point(shape = 22,
             color = "cyan4",
             fill = "yellow",
             size = 4,
             alpha = 0.4)
```

changing the aesthic based on different values of a variable

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length), alpha = 0.5) +
  theme_minimal()
```

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() + facet_wrap(~common_name, ncol = 3, scales = "free")
```

```{r}
lizards_counts <- lizards |> 
  group_by(common_name) |> 
  summarize(counts = n())
lizards_counts

liz_count <- lizards |> 
  count(common_name, site, tail)
liz_count
```

```{r}
ggplot(lizards_counts, aes(y = common_name, x = counts)) +
  geom_col() # default for categorical variable is alphabetical variable, can convert to factor and set the order
# convert and plot as factor by the value of n
# manual orders are not reproducible 
```

# Factor reorder

```{r}
ggplot(lizards_counts, aes(y = fct_reorder(common_name, counts), x = counts)) +
  geom_col()
```

Create scatterplots, faceted by species, for total length versus weight.

The color and shape of points should differ based on the species name (common name). Change theme. Make each faceted panel have flexible axis limits (scales should be free).

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name),
             alpha = 0.6,
             size = 3) +
  facet_wrap(~common_name, scales = "free") +
  theme_light() +
  labs(x = " Total Length (mm)",
       y = "Weight (grams)")
```

# Fun with `scales`

```{r}
ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "purple", high = "orange")

ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("magenta", "cyan4", "darkorchid3", "dodgerblue", "green", "yellow"))

ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "black")

ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "purple",mid = "white", high = "orange",
                     midpoint = 150,
                     breaks = c(50, 75,180,220,280))

ggplot(lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0 , to = 300, by = 60))
```

# Paletteer

Make a horizontal boxplot with common name on the y -axis, total length on the x- axis, with boxplot fill changing based on the common name

```{r}
# store updated variable as a factor to use whenever needed,f column, x reorder by
# put them in order by median of different levels of factor, default alphabetical,
# ordered by lowest to highest median
lizards_median <- lizards |> 
  mutate(common_name = fct_reorder(.f = common_name, .x = total_length, .fun = median))

ggplot(lizards, aes(y= common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name))

ggplot(data = lizards_median, aes(y= common_name, x = total_length)) + 
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "ggsci::default_gsea") + theme_minimal()


ggplot(data = lizards_median, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 500)) +
  scale_fill_paletteer_d(palette = "ggsci::light_blue_material") +
  labs(y = "Lizard species",
       x = "Total length (mm)") +
  theme_minimal()
palettes_d_names
```
# Customization through `theme()`
```{r}

# can use element_blank (), axis.title = will apply to both, highest resolution
# will apply
ggplot(data = lizards_median, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "red"),
        panel.grid.minor.y = element_line(color = "green"),
        axis.title.x = element_text(color = "purple", size = 10),
        axis.text.y = element_text(color = "green"),
        text = element_text(size = 18),
        panel.background = element_rect(color = 'purple', fill = 'yellow')) +
  annotate("text", x = 300, y = 50, label = "Wooo\nooo", color = "dodgerblue") +
  geom_vline(xintercept = 250, linetype = "dashed", color = "orange", size = 4)

```
# Repulsive labels with `ggrepel()`

```{r}
wws_lizards <- lizards |> 
  filter(common_name == "western whiptail", 
         site == "sand")
head(wws_lizards)

ggplot(data = wws_lizards, aes(x = total_length, y = weight)) +
  geom_point()+
  # geom_text(aes(label = toe_num))
  geom_text_repel(aes(label = toe_num))

head(gapminder)

gapminder |> 
  filter(year == 2002 & continent == "Europe") |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) + 
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)


ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point() +
  gghighlight(toe_num == 250, label_key = toe_num) # weight > 30
```
# Make a map
```{r}
# simple features 
jornada_veg <- read_sf(here::here("data_raw", "spatial_vegetation","doc.kml"))
# change polygons colors fill
ggplot() + geom_sf(data = jornada_veg,
                   aes(fill = Name),
                   color = NA) +
  theme_minimal() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

